---
title: "SV GWAS with AD/ADRD traits"
subtitle: "SVs with no LD with SNPs"
author:  |
 | Ricardo A. Vialle
 | Katia de Paiva Lopes
 |
 | Rush Alzheimer’s Disease Center
 | Chicago, IL 60612
date: "`r Sys.Date()`"
output: 
 rmarkdown::html_document:
   theme: spacelab
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true  
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(vcfR)
library(data.table)
library(ggrepel)
library(ggsci)
library(Biobase)
library(genefilter)
library(scales)
library(gt)

knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```

```{css zoom-lib-src, echo = FALSE, eval=F}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

```{r setup, include=FALSE}
root.dir = "~/projects/ADRD_SV_GWAS/"
runID = "run_20230209"

data.dir = paste0(root.dir,runID,"/")
figures.dir = paste0(data.dir,"/figures/")

load(paste0(data.dir,"ROSMAP_SV_4_association.RData"))
load(paste0(data.dir,"SV_pheno_list.RData"))
```

```{r Helper Functions, echo=FALSE}
source(paste0(data.dir,"/lib/plot_functions.R"))
source(paste0(data.dir,"/lib/support_functions.R"))
```

## Load data

```{r load_data, echo=FALSE}
res_ROS = fread(file = paste0(data.dir,"/","ROS_SVs","/association_results.tsv.gz"))
res_ROS %>% filter(pheno %in% rownames(pheno_data)) -> res_ROS
res_MAP = fread(file = paste0(data.dir,"/","MAP_SVs","/association_results.tsv.gz"))
res_MAP %>% filter(pheno %in% rownames(pheno_data)) -> res_MAP
res_META = fread(file = paste0(data.dir,"/Meta_analysis_results.tsv.gz")) %>% 
  filter(pheno %in% rownames(pheno_data))
res_ROSMAP = res_META %>% select(ID, sv_info, pheno, BETA_RE, STD_RE, PVALUE_RE2)
colnames(res_ROSMAP) = c("ID","sv_info","pheno","Estimate","Std. Error","pval")

ld_all = read.table(paste0("/pastel/resources/SVs/ld_sv_snv.tsv.gz"), header = T)
# Remove rows with self LD
ld_all2 = ld_all[!grepl("ALU|DEL|DUP|INS|LIN|SVA", ld_all$snv_id),]

gnomad_overlap = fread("/pastel/resources/SVs/rosmap_gnomad.rec_overlap_80")
gnomad_overlap$V5[grepl("ALU|SVA|LINE1", gnomad_overlap$V5)] = "INS" 
gnomad_overlap = gnomad_overlap[gnomad_overlap$V5 == gnomad_overlap$V11 | grepl("MCNV|CPX|BND",gnomad_overlap$V11), ]
```

## Prioritization

`PVALUE_RE2 <= 5e-3` and `MAF >= 0.05 & MAF <= 0.45` and `overlap with gnomAD-SV` and `no LD with SNPs`

```{r message=T}
svs_bestLD = ld_all2 %>% 
  group_by(sv_id) %>% 
  slice_max(n = 1, order_by = r2, na_rm = T, with_ties = F) %>% 
  mutate(LD = paste0(snv_id," (",r2,")"))

metasoft_output_noLD = res_META %>% 
  filter(!ID %in% svs_bestLD$sv_id) %>%
  mutate(MAF = as.numeric(gsub("(.*) maf:(.*))","\\2",sv_info))) %>%
  filter(MAF >= 0.05 & MAF <= 0.45) %>%
  filter(ID %in% unique(gnomad_overlap$V4)) %>%
  distinct() %>%
  arrange(PVALUE_RE2)

# dim(metasoft_output_noLD) # 3168   33
# length(unique(metasoft_output_noLD$ID)) # 132

ROSMAP_meta_hits = metasoft_output_noLD %>% 
  arrange(PVALUE_RE2) %>%
  dplyr::select(pheno, ID, PVALUE_RE2) %>% 
  filter(PVALUE_RE2 <= 5e-3)

ROSMAP_meta_hits$ID_pheno = paste0(ROSMAP_meta_hits$ID,"_",ROSMAP_meta_hits$pheno)

ROSMAP_meta_hits = ROSMAP_meta_hits %>% left_join(svs_bestLD[,c("sv_id","r2","LD")], by = c("ID" = "sv_id"))

message("Number of meta-analysis hits: ", nrow(ROSMAP_meta_hits))
message("Number of unique meta-analysis hits: ", length(unique(ROSMAP_meta_hits$ID)))
message("Number of unique meta-analysis hits with LD: ", length(unique(ROSMAP_meta_hits$ID[!is.na(ROSMAP_meta_hits$LD)])))
```

```{r}
res_ROSMAP = res_ROSMAP %>% 
  mutate(ID_pheno = paste0(ID,"_",pheno)) 

DiscROS = res_ROS %>% 
  mutate(ID_pheno = paste0(ID,"_",pheno)) %>%
  filter(ID_pheno %in% ROSMAP_meta_hits$ID_pheno)
DiscMAP = res_MAP %>%
  mutate(ID_pheno = paste0(ID,"_",pheno)) %>%
  filter(ID_pheno %in% ROSMAP_meta_hits$ID_pheno)

replROSMAP = DiscROS[,c("ID_pheno","ID","pheno","sv_info","closest_gene","Estimate","Std. Error","pval")] %>% 
  inner_join(DiscMAP[,c("ID_pheno","ID","pheno","Estimate","Std. Error","pval")], 
             by = c("ID_pheno","ID","pheno"), suffix = c(".ROS",".MAP")) %>%
  left_join(res_ROSMAP[,c("ID_pheno","ID","pheno","Estimate","Std. Error","pval")], by = c("ID_pheno","ID","pheno")) %>%
  left_join(ROSMAP_meta_hits[,c("ID_pheno","PVALUE_RE2","LD"),drop=F], by = c("ID_pheno")) %>% distinct()

res_ROSMAP$pheno_label = pheno_data[res_ROSMAP$pheno,"description"]
res_ROSMAP$pheno_variant = paste0(res_ROSMAP$pheno,":",res_ROSMAP$ID)

Results_table = data.frame(Phenotype = replROSMAP$pheno)
Results_table$ID = replROSMAP$ID
Results_table$SV = gsub(" maf.*)","bp)",replROSMAP$sv_info)
Results_table$MAF = percent(as.numeric(gsub("(.*) maf:(.*))","\\2",replROSMAP$sv_info)), accuracy = 0.1)
Results_table$BetaSE_ROS = sprintf("%.2f (%.2f)",replROSMAP$Estimate.ROS, replROSMAP$`Std. Error.ROS`)
Results_table$P_ROS = replROSMAP$pval.ROS
Results_table$BetaSE_MAP = sprintf("%.2f (%.2f)",replROSMAP$Estimate.MAP, replROSMAP$`Std. Error.MAP`)
Results_table$P_MAP = replROSMAP$pval.MAP
Results_table$BetaSE_ROSMAP = sprintf("%.2f (%.2f)",replROSMAP$Estimate, replROSMAP$`Std. Error`)
Results_table$P_ROSMAP = replROSMAP$PVALUE_RE2
Results_table$P_Meta = replROSMAP$PVALUE_RE2

Results_table$Gene = replROSMAP$closest_gene
Results_table$LD = replROSMAP$LD

Results_table$pheno_variant = paste0(Results_table$Phenotype,":",Results_table$ID)
res_ROSMAP$is2highlight = ifelse(res_ROSMAP$pheno_variant %in% Results_table$pheno_variant,"yes","no")

# fwrite(Results_table, file = paste0(data.dir,"meta_analysis_results",runSufix,".tsv.gz"), sep = "\t")

Results_table %>% arrange(P_ROSMAP) %>% 
  gt::gt() %>% fmt_scientific(columns = c("P_ROS","P_MAP","P_ROSMAP")) %>%
  tab_spanner( label = "ROS", columns = c(BetaSE_ROS, P_ROS)) %>%
  tab_spanner( label = "MAP", columns = c(BetaSE_MAP, P_MAP)) %>%
  tab_spanner( label = "Meta (ROS/MAP)", columns = c(BetaSE_ROSMAP, P_ROSMAP)) %>%
  tab_style(style = list(cell_text(style = "italic")), 
            locations = cells_body(columns = Gene)) %>%
  tab_header(
    title = "",
    subtitle = "Top results of genome-wide scan for SVs"
  )
```


```{r, fig.width=14, fig.height=6, eval=T}
p = plot_multitrait_chicago(res_final = res_ROSMAP, 
                               suggestive_threshold = 4, 
                               ylims = c(-8.5, 8.5), 
                               topN = nrow(res_ROSMAP), 
                               pheno_label = "pheno_label", 
                               is2highlight = "is2highlight")
p
```

```{r fig.width=10, fig.height=3.8}
selected_res = res_ROSMAP %>% 
  filter(ID %in% replROSMAP$ID) %>% 
  mutate(pheno_variant = paste0(pheno,":",ID)) %>% 
  arrange(pval) %>% distinct() 

selected_res_pval_mtx = selected_res %>% 
  dplyr::select(c(sv_info, pval, pheno_label)) %>%
  pivot_wider(id_cols = sv_info, names_from = pheno_label, values_from = pval) %>% 
  column_to_rownames("sv_info")
rownames(selected_res_pval_mtx) = gsub("(.*) maf:(.*))","\\1bp)",rownames(selected_res_pval_mtx))
selected_res_pval_mtx = selected_res_pval_mtx[gtools::mixedorder(gsub("(.*) (.*) (.*)","\\2",rownames(selected_res_pval_mtx))),]

log_matrix_pvalue_to_plot = -log10(selected_res_pval_mtx)
dimnames(log_matrix_pvalue_to_plot) = dimnames(log_matrix_pvalue_to_plot)

selected_res_pval_bonf_df = selected_res_pval_mtx %>% 
  apply(2, function(x) p.adjust(x, method = "bonferroni")) %>% 
  as.data.frame() 

selected_res_pval_bonf_df = selected_res_pval_mtx %>% 
  mutate(across(dplyr::everything(), ~ p.adjust(.x, method = "bonferroni", n = length(.x)))) %>% 
  as.data.frame() 

cut_not_signif = rowSums(selected_res_pval_bonf_df<0.05, na.rm = T)==0
selected_res_pval_bonf_df = selected_res_pval_bonf_df[!cut_not_signif,]
log_matrix_pvalue_to_plot = log_matrix_pvalue_to_plot[!cut_not_signif,]
selected_res_pval_mtx = selected_res_pval_mtx[!cut_not_signif,]

matrix_pvalue_to_plot = selected_res_pval_bonf_df
matrix_pvalue_to_plot.signif <- symnum(as.matrix(matrix_pvalue_to_plot), corr = FALSE, na = FALSE, 
                                           cutpoints = c(0, 0.05, 1), 
                                           symbols = c(quote("\u2731")," "))
library(ComplexHeatmap)
library(circlize)
library(RColorBrewer)

column_annotation = data.frame(pheno_label = colnames(matrix_pvalue_to_plot))
column_annotation %>% left_join(unique(res_ROSMAP[,c("pheno_label","pheno")])) %>%
  left_join(pheno_data %>% rownames_to_column("pheno")) %>%
  dplyr::select(category, pheno_label) %>%
  column_to_rownames("pheno_label") -> column_annotation
col_pal = RColorBrewer::brewer.pal(n = length(unique(column_annotation$category)), name ="Dark2")
names(col_pal) = sort(unique(column_annotation$category))

column_order = order(column_annotation$category)
column_annotation = column_annotation[column_order,,drop=F]
col_ha = columnAnnotation(`Category` = column_annotation$category,
                       col = list(`Category` = col_pal),
                       show_legend = c(F), 
                       height = unit(0.3,"mm"), 
                       annotation_height = unit(0.3, "mm"), 
                       simple_anno_size = unit(2, "mm"),
                       annotation_label = "")

##### Show Overlaps
selected_res = add_SV_annot(as.data.frame(selected_res))

row_annotation = data.frame(sv_info = rownames(matrix_pvalue_to_plot))
rownames(row_annotation) = row_annotation$sv_info
row_annotation = row_annotation %>% 
  left_join(selected_res %>% 
              mutate(sv_info = gsub("(.*) maf:(.*))","\\1bp)",sv_info)) %>%
              dplyr::select(sv_info, overlapGene, overlapIntronic, overlapIntergenic) %>%
              distinct() %>% pivot_longer(-sv_info, names_to = "Annotation") %>% filter(value == T), 
            by = "sv_info")
row_annotation$Annotation = gsub("overlap","",row_annotation$Annotation)
row_annotation$Annotation = gsub("Gene","Genic",row_annotation$Annotation)

row_pal = wesanderson::wes_palette("Darjeeling1")[1:3]
row_pal = viridis_pal(option = "F")(3)
names(row_pal) = c("Genic", "Intronic", "Intergenic")

row_ha = rowAnnotation(
  ` ` = row_annotation$Annotation,
  col = list(` ` = row_pal),
  show_legend = T,
  width = unit(0.3,"mm"), 
  annotation_width = unit(0.3, "mm"), 
  simple_anno_size = unit(2, "mm"))

log_matrix_pvalue_to_plot = log_matrix_pvalue_to_plot[,column_order]
matrix_pvalue_to_plot.signif = matrix_pvalue_to_plot.signif[,column_order]

Heatmap(as.matrix(log_matrix_pvalue_to_plot), 
        name = "-log10(P-value)",
        cell_fun = function(j, i, x, y, width, height, fill) {
            grid.text( matrix_pvalue_to_plot.signif[i,j], x, y, gp = gpar(fontsize = 12))
        },
        col = colorRampPalette((brewer.pal(n = 7, name ="Reds")))(100),
        row_names_side = "left", show_row_names = T,
        cluster_rows = F, cluster_columns = F,
        column_names_rot = 90,
        column_names_gp = gpar(fontsize = 9),
        row_names_gp = gpar(fontsize = 9),
        bottom_annotation = col_ha,
        left_annotation = row_ha,
        show_row_dend = F, show_column_dend = F, 
        rect_gp = gpar(col = "white", lwd = 1))
```

## Session info

```{r}
sessionInfo()
```
